---
title: 'Data Wrangling: Dealing With Missing Data'
author: "Jonathan Freeland"
date: "March 15, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
library(tidyr)
library(mice)
```
#### Import Data
```{r, cache=TRUE}
data <- read_csv("titanic_original.csv")
```

#### Handle Missing Values
##### Embarked
There are `r NROW(which(is.na(data$embarked)))` missing `embarked` value(s) which are known to be Southampton, replace with `S`.
```{r}
data$embarked[is.na(data$embarked)] <- "S"
```

##### Age
There are `r NROW(which(is.na(data$age)))` missing `age` value(s). Replace with overall mean, or try to be more accurate with `mice`?
```{r, cache=TRUE}
missing_ages = is.na(data$age)
# simple imputation with mean, skip this
##data$age[is.na(data$age)] <- na.omit(data$age) %>% mean()
##data$age[is.na(data$age)] <- with(data, impute(age, mean))

# imputation with mice
age_factors <- c("pclass","sex","embarked","sibsp","parch")
data[age_factors] <- lapply(data[age_factors], function(x) as.factor(x))
mice_mod <- mice(data[, names(data) %in% c(age_factors,"age")], method="rf", printFlag=FALSE)
mice_output <- complete(mice_mod)

# visualize if imputation closely matches original data
par(mfrow=c(1,2))
hist(data$age, freq=F, main="age: original", col="darkblue", ylim=c(0,0.04))
hist(mice_output$age, freq=F, main="age: mice", col="lightblue", ylim=c(0,0.04))

# replace missing values with imputed ones
data$age[missing_ages] <- mice_output$age[missing_ages]
```

##### Boat
There are `r NROW(which(is.na(data$boat)))` missing `boat` values(s), replace with `-1`.
```{r}
data$boat[is.na(data$boat)] <- -1
```

##### Cabin
Add `has_cabin_number` column indicating whether passenger has a cabin. What is survival rate of people with cabins versus without?
```{r}
data$has_cabin_number <- as.numeric(!is.na(data$cabin))
cabin_survival = NROW(data %>% filter(has_cabin_number == 1, survived == 1)) / NROW(data %>% filter(has_cabin_number == 1))
non_cabin_survival = NROW(data %>% filter(has_cabin_number == 0, survived == 1)) / NROW(data %>% filter(has_cabin_number == 0))
```
Cabin surival rate: `r cabin_survival*100`%; non-cabin survival rate: `r non_cabin_survival*100`%.

#### Export Data
```{r}
write.csv(data, "titanic_clean.csv")
```